Poincar\'e--Friedrichs constants over triangulated domains

### 2024-06-11 main.tex

We address the construction of potentials for the gradient, the curl and the divergence operators, and more generally the exterior derivative, over triangulated domains. The potential operators are ought to satisfy bounds in Lebesgue norms that are explicitly computable in terms of the geometry. 
In particular, this yields computable upper bounds for Poincar\'e--Friedrichs inequalities. 
For example, if $\bff \in \bfL^{p}(\Omega)$ is a square-integrable vector field over some domain $\Omega \subseteq \bbR^3$, 
we want to bound the $L^p$ norm of the potential
\begin{align*}
    \argmin\limits_{ \substack{ \bfv \in \bfL^{p}(\curl) \\ \curl \bfv = \bff } } \| \bfv \|_{L^{p}(\Omega)}
    \leq 
    C_{\curl}
    \| \bff \|_{L^{p}(\Omega)}
    .
\end{align*}
Finding a potential operator, whether it be linear or nonlinear, and bounding the Poincar\'e--Friedrichs constant is generally not an easy undertaking. 
We address this topic in the special case where $\Omega$ admits a shellable triangulation.
While this constraints us to the class of {\color{red}contractible} domains, 
having computable upper bounds for such domains is an important step stone towards the more general case. 
For example, local patches around a simplex within a larger triangulation belong to the class of shellable triangulations.

@@

\begin{remark}
    Our use of the term \textit{face} as is common in the finite element literature
    and is synonymous with \textit{facet} as used in the literature on polyhedral combinatorics.
    Notably, this terminology difers from the uses \textit{face} and \textit{facet} in the theory of polyhedra. 
\end{remark}

@@

We now gather a few helpful properties of simplicial manifolds,
which we state without proof.

\begin{lemma}\label{lemma:characterizationofmanifoldcomplexes}
    Let $\calT$ be a finite $n$-dimensional simplicial manifold. 
    \begin{enumerate}
        \item Any face $F \in \Faces(\calT)$ is not contained in the boundary if and only if it is contained in exactly two $n$-simplices.
        \item Any face $F \in \Faces(\calT)$ is contained in the boundary if and only if it is contained in exactly one $n$-simplex.
        % \item $\calT$ is non-branching, that is, every $F \in \calT$ of dimension $n-1$ is contained in at most two $n$-simplices of $\calT$. 
        \item If $S$ is an inner simplex, then $\patch_{\calT}(S)$ is a simplicial $n$-ball with $S$ as an inner simplex
        and $\carapace_{\calT}(S)$ is a simplicial $(n-1)$-sphere. 
        \item If $S$ is a boundary subsimplex, then $\patch_{\calT}(S)$ is a simplicial $n$-ball with $S$ as a boundary simplex,
        and $\carapace_{\calT}(S)$ is a simplicial $(n-1)$-ball.
        \item If the underlying space of $\calT$ is connected, then $\calT$ is strongly connected.
    \end{enumerate}
\end{lemma}
\begin{proof}
    % Faces have at most two parents TODO must be more careful here 
    Let $F \in \Faces(T)$ and let $z_F \in F$ be its midpoint. 
    Let $B_F$ be a ball around $z_F$ so small that it only intersects those $n$-simplices of $\calT$ that already contain $z_F$.
    Suppose there were distinct $n$-simplices $T_1, T_2, T_3$ that contain $z_F$. The intersection of any two of them is $F$, and their interiors are disjoint. 
    It follows that $B_F \setminus F$ has at least $3$ connected components, though $B_F \setminus F$ has at most two components. 
    We conclude that each face is in at most two $n$-simplices.

    % TODO inner/boundary faces

    % TODO structure of stars around simplices

    % TODO

    % TODO stars are strongly connected 
    We show that each vertex star is strongly connected via a short induction argument.
    Clearly, any simplicial $1$-ball and simplicial $1$-sphere is strongly connected. 
    Now, if $n \geq 1$, then any vertex star in an $(n+1)$-dimensional simplicial manifold is strongly connected if all simplicial $n$-balls and $n$-spheres are strongly connected.
    The induction argument implies that each vertex star in $\calT$ is strongly connected.

    % strongly connected
    If the underlying space is connected, then we easily see that the underlying space of the $1$-skeleton is path-connected.
    Given $n$-simplices $S,S' \in \calT$, 
    we can thus choose a sequence of $\hat S_0=S,\hat S_1,\dots,\hat S_m=S' \in \calT$ such that $\hat S_{i} \cap \hat S_{i-1} \neq \emptyset$ for all $1 \leq i \leq m$.
    We also notice that each vertex star is strongly connected. 
    We can thus construct change the first sequence to another $S_0=S,S_1,\dots,S'=S_m \in \calT$ 
    such that $S_{i} \cap S_{i-1}$ is a face of both $S_{i}$ and $S_{i-1}$ for all $1 \leq i \leq m$.
\end{proof}

@@

\begin{lemma}
    Any simplicial $2$-disk is shellable.
    Any simplicial $2$-sphere is shellable. 
\end{lemma}
\begin{proof}
    Let $\calT$ be a simplicial $2$-disk. 
    If $\calT$ contains only one triangle, then there is nothing to show. 
    Otherwise, there exists at least one triangle $T$ that touches the boundary of the disk at either one or two edges {\color{red}Needs proof}. 
    Letting $\calT'$ be the triangulation obtained by removing $T$,
    we have another triangulation of the $2$-disk that intersects $T$ only at either two or one edges.
    Repeating this procedure yields an enumeration of the triangles of $\calT$,
    whose reverse is a shelling.

@@

Suppose that $\calT$ is a strongly connected $n$-dimensional simplicial complex that triangulates a manifold, and suppose that its boundary complex $\partial\calT$ is either empty or strongly connected.  
Suppose furthermore that $T$ is an $n$-simplex, $T \notin \calT$, whose boundary complex intersects with $\partial\calT$ at a non-empty simplicial complex $\calS$ of dimension $n-1$. In particular, $\calS$ is either the whole boundary complex of $T$ or a patch around a proper subsimplex of $T$. We write $\calT'$ be the  simplicial complex that contains $\calT \cup T$.
We exclude that case $n \leq 1$ since otherwise the discussion is trivial.

According to the simplicial Mayer-Vietoris theorem, there exists an exact sequence 
\begin{align*}
    \begin{CD}
        \dots \longrightarrow H_{k}(\calS) \longrightarrow H_{k}(\calT \oplus T) \longrightarrow H_{k}(\calT') \longrightarrow H_{k-1}(\calS) \longrightarrow\dots 
    \end{CD}
\end{align*}
We notice that $\calS$, $\calT$, $T$ and $\calT'$ are each connected, and hence their zeroth homology groups are isomorphic to $\bbZ$. 
By the nature of the mappings in the Mayer-Vietoris sequence, 
\color{red}which uses some specific knowledge about the arrows\color{black}, 
we see that 
\begin{align*}
    \begin{CD}
        0 \longrightarrow H_{0}(\calS) \longrightarrow H_{0}(\calT) \oplus H_{0}(T) \longrightarrow H_{0}(\calT') \longrightarrow 0 
    \end{CD}
\end{align*}
is an exact sequence. 
Moreover, $H_{k}(\calS) = H_{k}(T) 0$ for $0 < k < n-1$. Hence, for $1 < k < n-1$,
the Mayer-Vietoris sequence includes the exact sequence 
\begin{align*}
    \begin{CD}
        0 \longrightarrow H_{k}(\calT \oplus T) \longrightarrow H_{k}(\calT') \longrightarrow 0 
    \end{CD}
\end{align*}
and we conclude that $H_{k}(\calT) = 0$ for $0 < k < n-1$.
Finally, the Mayer-Vietoris sequence includes the exact sequences 
\begin{align*}
    \begin{CD}
        0 \longrightarrow H_{n}(\calT \oplus T) \longrightarrow H_{n}(\calT') \longrightarrow H_{n-1}(\calS) \longrightarrow H_{n-1}(\calT \oplus T) \longrightarrow H_{n-1}(\calT') \longrightarrow 0 
    \end{CD}
\end{align*}
% Here, if $n=1$, it already follows that $H_{n}(\calT \oplus T) \simeq H_{n}(\calT')$.
% % We know that  $H_{0}(\calT') \simeq H_{0}(\calT) \simeq H_{0}(T) \simeq H_{0}(\calS) \simeq \bbZ$ because the respective simplicial complexes are path-connected. 
% Hence $H_{1}(\calT) \oplus H_{1}(T) \simeq H_{1}(\calT')$. 
% one easily sees that $H_{0}(\calT') \simeq \bbZ$ and $H_{1}(\calT) \oplus H_{1}(T) \simeq H_{1}(\calT')$. 
Since $\calT$ and $T$ are triangulations of manifolds with non-empty boundary, their $n$-th simplicial homology groups vanish. 
In the case that $\calS$ is isomorphic to disk,
then $H_{n}(\calT') \simeq H_{n}(\calT \oplus T)$ are trivial and $H_{n-1}(\calT') \simeq H_{n-1}(\calT \oplus T)$ follows.
Consider now the case that $\calS$ is isomorphic to a sphere.
Since the boundary complex of $\calT$ is strongly connected and non-branching, it must coincide with $\calS$. Hence $\calT'$ triangulates a manifold without boundary, and thus $H_{n}(\calT') \simeq \bbZ$. 
\color{red}By some hand-waving argument, $H_{n}(\calT') \longrightarrow H_{n-1}(\calS)$ is an isomorphism.\color{black}
Hence $H_{n-1}(\calT) \rightarrow H_{n-1}(\calT')$ is an isomorphism.

We conclude that the simplicial homology groups of a shellable $n$-dimensional simplicial complex $\calT$ that triangulates a manifold satisfy 
\begin{align*}
    H_0(\calT) \simeq \bbZ, 
    \quad 
    H_1(\calT) \simeq \dots \simeq H_{n-1}(\calT) \simeq 0.
\end{align*}
Moreover, $H_n(\calT) \simeq \bbZ$ if $\calT$ has no boundary and $H_n(\calT) \simeq 0$ otherwise.

@@

@@

\section{Constructive estimate of Poincar\'e--Friedrichs constants}

>>>>

We address the construction of potentials for the gradient, the curl and the divergence operators, and more generally the exterior derivative, over triangulated domains. The potential operators are ought to satisfy bounds in Lebesgue norms that are explicitly computable in terms of the geometry. 
In particular, this yields computable upper bounds for Poincar\'e--Friedrichs inequalities. 
For example, if $\bff \in \bfL^{2}(\Omega)$ is a square-integrable vector field over some domain $\Omega \subseteq \bbR^3$, 
we want to bound the $\bfL^2$ norm of the potential
\begin{align*}
    \argmin\limits_{ \substack{ \bfv \in \bfL^{2} \\ \curl \bfv = \bff } } \| \bfv \|_{\bfL^{2}(\Omega)}
    \leq 
    C_{\curl}
    \| \bff \|_{\bfL^{2}(\Omega)}
    .
\end{align*}
Finding a potential operator, whether it be linear or nonlinear, and bounding the Poincar\'e--Friedrichs constant is generally not an easy undertaking. 
We address this topic in the special case where $\Omega$ admits a shellable triangulation.
Even though only contractible domains can admit a shellable triangulation, 
having computable upper bounds for such domains is an important step stone towards the more general case. 
For example, local patches around a simplex within a larger triangulation belong to the class of shellable triangulations.

We take the Poincar\'e--Friedrichs inequality for the gradient of scalar functions as a conceptual point of reference.
Given a domain $\Omega \subseteq \bbR^n$ and $1 \leq p \leq \infty$,
we search for a constant $C > 0$ such that 
\begin{align*}
    \min\limits_{c\in\bbR}
    \| u - c \|_{L^{p}(\Omega)}
    \leq 
    C_{\grad,p} \| \nabla u \|_{L^p(\Omega)},
    \quad 
    u \in W^{1,p}(\Omega)
    .
\end{align*}
Finding a potential of a given gradient $\nabla u$ is generally a nonlinear operation. 
The operator norm of any \emph{linear} potential construction is an upper bound for the Poincar\'e--Friedrichs constant. 
Even though the existence of such a bounded linear potential construction is not evident a priori, 
picking the average-free potential 
\begin{align*}
    \Phi_{\grad}( \bff ) 
    := 
    \argmin\limits_{ \substack{ v \in W^{1,p}(\Omega) \\ \int_\Omega v = 0 } } \| \nabla v - \bff \|_{L^{p}(\Omega)},
    \quad 
    \bff \in \nabla W^{1,p}(\Omega)
\end{align*}
is a linear operation. Its operator norm is the constant in the Poincar\'e inequality 
\begin{align*}
    \| u - u_{\Omega} \|_{L^{p}(\Omega)}
    \leq 
    C_{\rm P} \| \nabla u \|_{L^p(\Omega)},
    \quad 
    u \in W^{1,p}(\Omega)
    ,
\end{align*}
where $u_\Omega$ denotes the average of $u$.
Considerable research effort has gone into finding (or estimating) the optimal constants in the Poincar\'e--Friedrichs inequalities or Poincar\'e inequalities.

The situation is less clear cut when we study the curl or divergence operators
\begin{align*}
    \curl : \bfW^{p}(\curl,\Omega) \rightarrow \bfL^{p}(\curl,\Omega),
    \qquad 
    \divergence : \bfW^{p}(\divergence,\Omega) \rightarrow L^{p}(\Omega),
\end{align*}
where we use the function spaces
\begin{align*}
    \bfW^{p}(\curl,\Omega) &:= \left\{ \bfu \in \bfL^p(\Omega) : \curl \bfu \in \bfL^p(\Omega) \right\}
    ,
    \\
    \bfW^{p}(\divergence,\Omega) &:= \left\{ \bfu \in \bfL^p(\Omega) : \divergence \bfu \in L^p(\Omega) \right\}
    .
\end{align*}
The fundamental difference is that the curl and divergence have infinite-dimensional kernels. 
{\color{red}The existence of a bounded linear potential operator is a non-trivial fact even in the Hilbert space case. }
In the Banach space case, it is not immediately evident that the kernels of these operators are complemented.
We are interested in constants for inequalities such as 
\begin{align*}
    \argmin\limits_{ \substack{ v \in W^{p}(\curl,\Omega) \\ \curl v = \bff } } \| v \|_{L^{p}(\Omega)}
    \leq 
    C_{\curl}
    \| \bff \|_{L^{p}(\Omega)}
    ,
    \qquad 
    \argmin\limits_{ \substack{ v \in W^{p}(\divergence,\Omega) \\ \divergence v = f } } \| v \|_{L^{p}(\Omega)}
    \leq 
    C_{\divergence}
    \| f \|_{L^{p}(\Omega)}
    ,
\end{align*}
and whether these constants can be estimated by the operator norm of linear potential operators 
\begin{align*}
    \Phi_{\curl} : \curl \bfW^{p}(\curl,\Omega) \rightarrow \bfW^{p}(\curl,\Omega)
    ,
    \quad 
    \Phi_{\divergence} : \divergence \bfW^{p}(\divergence,\Omega) \rightarrow \bfW^{p}(\divergence,\Omega)
    .
\end{align*}
We remark that seemingly not much attention has been given to the study of optimal constants in such Poincar\'e--Friedrichs inequalities for the curl and divergence operators, 
% even if the domain is convex, 
which is in sharp contrast to the study of the gradient operator.

We explore formulate our main results in the framework of exterior calculus.
The formalism of exterior calculus generalizes vector calculus to arbitrary dimensions and allows us to study many differential operators, such as the curl and divergence, in a unified manner. 
%
It is used ubiquitously in physics and engineering. 
%
For the purpose of our discussion, this formalisms allows us to leverage results from a larger body of literature in differential geometry and functional analysis. 
\\

Before we outline the main results of this exposition in more details, 
we first review the state of the literature on Poincar\'e--Friedrichs inequalities on specific classes of domains.

% Convex domains

% Domains starshaped wrt ball

% Limitations

Most research on optimal constants has addressed the case of convex domains. 
Here, the optimal constants, in terms of the domain diameter alone, are known if $p=2$ \cite{bebendorf2003note} and if $p=1$ \cite{acosta2004optimal}. 
Notably, the optimal constants $C_{\grad,p}$ for all $1 \leq p < \infty$, again in terms of only the domain diameter, are known when the domain is convex.~\cite{esposito2013poincare,ferone2012remark}.

In the Hilbert space setting when $p = 2$, Guerini and Savo~\cite{guerini2004eigenvalue} have develop lower bounds for the spectrum of the Hodge--Laplace operator, which implies upper bounds for the Poincar\'e--Friedrichs constant of the exterior derivative, and hence for the curl and divergence operators.
\\

@@

\begin{remark}
    Our use of the term \textit{face} as is common in the finite element literature
    and is synonymous with \textit{facet} as used in the literature on polyhedral combinatorics.
    Notably, this terminology differs from the uses \textit{face} and \textit{facet} in the theory of polyhedra. 
\end{remark}

@@

We gather a few helpful properties of simplicial manifolds.

\begin{lemma}\label{lemma:characterizationofmanifoldcomplexes}
    Let $\calT$ be a finite $n$-dimensional simplicial manifold. 
    \begin{itemize}
        \item Any face $F \in \Faces(\calT)$ is not contained in the boundary if and only if it is contained in exactly two $n$-simplices.
        \item Any face $F \in \Faces(\calT)$ is contained in the boundary if and only if it is contained in exactly one $n$-simplex.
        \item The simplices contained in the boundary constitute a triangulation of the boundary.
        % \item $\calT$ is non-branching, that is, every $F \in \calT$ of dimension $n-1$ is contained in at most two $n$-simplices of $\calT$. 
        \item If $S$ is an inner simplex, then $\patch_{\calT}(S)$ is a simplicial $n$-ball with $S$ as an inner simplex
        and $\carapace_{\calT}(S)$ is a simplicial $(n-1)$-sphere. 
        \item If $S$ is a boundary subsimplex, then $\patch_{\calT}(S)$ is a simplicial $n$-ball with $S$ as a boundary simplex,
        and $\carapace_{\calT}(S)$ is a simplicial $(n-1)$-ball.
        \item If the underlying space of $\calT$ is connected, then $\calT$ is strongly connected.
    \end{itemize}
\end{lemma}
\begin{proof}
    \begin{enumerate}
    \item 
    We will use the following fact: if $S \in \calT$ has an inner point that lies on $\partial M$, then all inner points of $S$ are on $\partial M$. 
    Since the boundary is closed, every $S \in \calT$ is either a subset of the boundary or all its inner points lie in the interior of the manifold.

    \item
    Let $F \in \Faces(T)$ and let $z_F \in F$ be its midpoint. 
    Let $\mathring B_F$ be an open neighborhood around $z_F$ so small that it only intersects those $n$-simplices of $\calT$ that already contain $z_F$ and no faces other than $F$.
    Suppose there are distinct $n$-simplices $T_1, T_2, \dots, T_K$ that contain $z_F$. 
    The intersection of any two of them is $F$, and their interiors are disjoint.

    If $z_F$ is an interior point of $M$, 
    then we can assume that $\mathring B_F$ is homeomorphic to an open $n$-ball and $\partial \mathring B_F$ is homeomorphic to a sphere $S$ of dimension $n-1$. 
    At the same time, $X = F \subseteq \partial \mathring B_F$ is empty if $n=1$ or homeomorphic to a sphere $S' \subseteq S$ of dimension $n-2$. 
    We see that $\partial \mathring B_F \setminus X$ has $K$ distinct connected components. 
    But by construction $\partial \mathring B_F \setminus X$ is homeomorphic to $S \setminus S'$,
    which has exactly two distinct connected components according to the Jordan-Brouwer separation theorem.
    We conclude that $F$ is contained in two $n$-simplices of $\calT$.
    If an inner face $F$ were contained in only one $n$-simplex,
    then we could construct a homeomorphism from the $n$-ball onto itself that maps an interior point on the its boundary, which cannot be.
    So each inner face is contained in exactly two $n$-simplices.

    If $F$ is contained in only one $n$-simplex, then it must be a boundary face. 
    If a boundary face $F$ were contained in at least two $n$-simplices,
    then we could construct a homeomorphism from the $n$-ball onto itself that maps an interior point on the its boundary, which cannot be.
    So each boundary face is contained in exactly one $n$-simplex.

    \item
    Clearly, the simplices of $\calT$ contained in the boundary constitute a simplicial complex. 
    If $x \in \partial M$ is a boundary simplex, then it is an inner point of some simplex $S \in \calT$.
    But then this simplex must be a boundary simplex, so the boundary simplices form a triangulation of $\partial M$.

    \item 
    % TODO structure of stars around simplices 
    Let $S$ be any simplex with midpoint $z_F$. 
    For $\delta > 0$ small enough, 
    \begin{align*}
        B_{\delta}(z_F) = \left\{ x \in M \suchthat d(x,z_F) \leq \delta \right\},
        \quad 
        S_{\delta}(z_F) = \left\{ x \in M \suchthat d(x,z_F) = \delta \right\},
    \end{align*}
    are subsets of $\patch_{\calT}(S)$.
%     For any $K \in \patch_{\calT}(S)$ that contains $S$, 
%     we can easily find a homeomorphism

%     Then $\patch_{\calT}(S)$ is star-shaped with respect to $z_F$. 
%     We let $B$
%     Any half-ray starting at $z_F$ intersects at most one vertex in $\patch_{\calT}(S)$ other than $z_F$. 
%

    % TODO stars are strongly connected 
    \item 
    We show that each vertex star is strongly connected via a short induction argument.
    Clearly, any simplicial $1$-ball and simplicial $1$-sphere is strongly connected. 
    Now, if $n \geq 1$, then any vertex star in an $(n+1)$-dimensional simplicial manifold is strongly connected if all simplicial $n$-balls and $n$-spheres are strongly connected.
    The induction argument implies that each vertex star in $\calT$ is strongly connected.

    % strongly connected
    \item 
    If the underlying space is connected, then we easily see that the underlying space of the $1$-skeleton is path-connected.
    Given $n$-simplices $S,S' \in \calT$, 
    we can thus choose a sequence of $\hat S_0=S,\hat S_1,\dots,\hat S_m=S' \in \calT$ such that $\hat S_{i} \cap \hat S_{i-1} \neq \emptyset$ for all $1 \leq i \leq m$.
    We also notice that each vertex star is strongly connected. 
    We can thus construct change the first sequence to another $S_0=S,S_1,\dots,S'=S_m \in \calT$ 
    such that $S_{i} \cap S_{i-1}$ is a face of both $S_{i}$ and $S_{i-1}$ for all $1 \leq i \leq m$.
    \end{enumerate}
\end{proof}

@@

\begin{lemma}    
    If an $n$-dimensional simplicial complex $\calT$ has a shelling $T_{0}, T_{1}, T_{2}, \dots$ and is non-branching, then it triangulates a manifold, possibly with boundary.
    In particular, $\patch_{\calT}(T_{0}) \cup \patch_{\calT}(T_{1}) \cup \dots \cup \patch_{\calT}(T_{m})$ triangulates a manifold for all $m \geq 0$.
\end{lemma}
\begin{proof}    
\end{proof}

The notion of shelling is taken from Ziegler's book~\cite{ziegler2012lectures}
\begin{lemma}    
    Any simplicial $2$-disk is shellable.
    Any simplicial $2$-sphere is shellable. 
\end{lemma}
\begin{proof}
    Let $\calT$ be a simplicial $2$-disk. 
    If $\calT$ contains only one triangle, then there is nothing to show. 
    Otherwise, there exists at least one triangle $T$ that touches the boundary of the disk at either one or two edges {\color{red}Needs proof}. 
    Letting $\calT'$ be the triangulation obtained by removing $T$,
    we have another triangulation of the $2$-disk that intersects $T$ only at either two or one edges.
    Repeating this procedure yields an enumeration of the triangles of $\calT$,
    whose reverse is a shelling.

@@

    Suppose furthermore that $T$ is an $n$-simplex, $T \notin \calT$, 
    whose boundary complex intersects with $\partial\calT$ at a non-empty simplicial complex $\calS$ of dimension $n-1$. 
    In particular, $\calS$ is either the whole boundary complex of $T$ or a patch around a proper subsimplex of $T$. 
    We write $\calT'$ for the  simplicial complex that contains $\calT \cup T$.

Suppose that $\calT$ is a strongly connected $n$-dimensional simplicial complex that triangulates a manifold, and suppose that its boundary complex $\partial\calT$ is either empty or strongly connected.  
Suppose furthermore that $T$ is an $n$-simplex, $T \notin \calT$, whose boundary complex intersects with $\partial\calT$ at a non-empty simplicial complex $\calS$ of dimension $n-1$. In particular, $\calS$ is either the whole boundary complex of $T$ or a patch around a proper subsimplex of $T$. We write $\calT'$ be the  simplicial complex that contains $\calT \cup T$.
We exclude the case $n \leq 1$ since otherwise the discussion is trivial.

According to the simplicial Mayer-Vietoris theorem, there exists an exact sequence 
\begin{align*}
    \begin{CD}
        \dots \longrightarrow H_{k}(\calS) \longrightarrow H_{k}(\calT \oplus T) \longrightarrow H_{k}(\calT') \longrightarrow H_{k-1}(\calS) \longrightarrow\dots 
    \end{CD}
\end{align*}
We notice that $\calS$, $\calT$, $T$ and $\calT'$ are each connected, and hence their zeroth homology groups are isomorphic to ${G}$. 
By the nature of the mappings in the Mayer-Vietoris sequence, 
\color{red}which uses some specific knowledge about the arrows\color{black}, 
we see that 
\begin{align*}
    \begin{CD}
        0 \longrightarrow H_{0}(\calS) \longrightarrow H_{0}(\calT) \oplus H_{0}(T) \longrightarrow H_{0}(\calT') \longrightarrow 0 
    \end{CD}
\end{align*}
is an exact sequence. 
Moreover, $H_{k}(\calS) = H_{k}(T) = 0$ for $0 < k < n-1$. Hence, for $1 < k < n-1$,
the Mayer-Vietoris sequence includes the exact sequence 
\begin{align*}
    \begin{CD}
        0 \longrightarrow H_{k}(\calT \oplus T) \longrightarrow H_{k}(\calT') \longrightarrow 0 
    \end{CD}
\end{align*}
and we conclude that $H_{k}(\calT) = 0$ for $0 < k < n-1$.
Finally, the Mayer-Vietoris sequence includes the exact sequence 
\begin{align*}
    \begin{CD}
        0 \longrightarrow H_{n}(\calT \oplus T) \longrightarrow H_{n}(\calT') \longrightarrow H_{n-1}(\calS) \longrightarrow H_{n-1}(\calT \oplus T) \longrightarrow H_{n-1}(\calT') \longrightarrow 0 
    \end{CD}
\end{align*}

% Here, if $n=1$, it already follows that $H_{n}(\calT \oplus T) \simeq H_{n}(\calT')$.
% % We know that  $H_{0}(\calT') \simeq H_{0}(\calT) \simeq H_{0}(T) \simeq H_{0}(\calS) \simeq \bbZ$ because the respective simplicial complexes are path-connected. 
% Hence $H_{1}(\calT) \oplus H_{1}(T) \simeq H_{1}(\calT')$. 
% one easily sees that $H_{0}(\calT') \simeq \bbZ$ and $H_{1}(\calT) \oplus H_{1}(T) \simeq H_{1}(\calT')$. 
Since $\calT$ and $T$ are triangulations of manifolds with non-empty boundary, their $n$-th simplicial homology groups vanish. 
then $H_{n}(\calT') \simeq H_{n}(\calT \oplus T)$ are trivial and $H_{n-1}(\calT') \simeq H_{n-1}(\calT \oplus T)$ follows.
Consider now the case that $\calS$ is isomorphic to a sphere.
Since the boundary complex of $\calT$ is strongly connected and non-branching, it must coincide with $\calS$. Hence $\calT'$ triangulates a manifold without boundary, and thus $H_{n}(\calT') \simeq \bbZ$. 
\color{red}By some hand-waving argument, $H_{n}(\calT') \longrightarrow H_{n-1}(\calS)$ is an isomorphism.\color{black}
Hence $H_{n-1}(\calT) \rightarrow H_{n-1}(\calT')$ is an isomorphism.

In the case that $\calS$ is isomorphic to a disk, we have $H_{n-1}(\calS) = 0$,
and then $H_{n}(\calT') \simeq H_{n}(\calT \oplus T) = 0$ are trivial 
and $H_{n-1}(\calT') \simeq H_{n-1}(\calT \oplus T)$ follows.
Consider now the case that $\calS$ is isomorphic to a sphere.
Since the boundary complex of $\calT$ is strongly connected and non-branching, it must coincide with $\calS$. Hence $\calT'$ triangulates a manifold without boundary, and thus $H_{n}(\calT') \simeq {G}$. 
\color{red}By some hand-waving argument, $H_{n}(\calT') \longrightarrow H_{n-1}(\calS)$ is an isomorphism.\color{black}
Hence $H_{n-1}(\calT) \rightarrow H_{n-1}(\calT')$ is an isomorphism.

We conclude that the simplicial homology groups of a shellable $n$-dimensional simplicial complex $\calT$ that triangulates a manifold satisfy 
\begin{align*}
    H_0(\calT) \simeq \bbZ, 
    \quad 
    H_1(\calT) \simeq \dots \simeq H_{n-1}(\calT) \simeq 0.
\end{align*}
Moreover, $H_n(\calT) \simeq {G}$ if $\calT$ has no boundary and $H_n(\calT) \simeq 0$ otherwise. 
Moreover, $H_n(\calT) \simeq \bbZ$ if $\calT$ has no boundary and $H_n(\calT) \simeq 0$ otherwise.

@@

\begin{remark}    
    % % We know that  $H_{0}(\calT') \simeq H_{0}(\calT) \simeq H_{0}(T) \simeq H_{0}(\calS) \simeq {G}$ because the respective simplicial complexes are path-connected. 
    % Hence $H_{1}(\calT) \oplus H_{1}(T) \simeq H_{1}(\calT')$. 
    % one easily sees that $H_{0}(\calT') \simeq {G}$ and $H_{1}(\calT) \oplus H_{1}(T) \simeq H_{1}(\calT')$.

    % then $H_{n}(\calT') \simeq H_{n}(\calT \oplus T)$ and $H_{n-1}(\calT') \simeq H_{n-1}(\calT \oplus T)$ follow immediately.
    % Moreover, Thus $H_{n}(\calT') = H_{n}(\calT) \oplus H_{n}(\calS)$.
    % We conclude that any shellable simplicial complex has the homology groups of a disk or a sphere. 
    % NB: the zeroth homology is always free, and the n-th homology group, being the kernel of a mapping free abelian group, is free as well.
    % \begin{align*}
    %     \begin{CD}
    %         0 \longrightarrow H_{1}(\calT) \oplus H_{1}(T) \longrightarrow H_{1}(\calT') \longrightarrow H_{0}(\calS) \longrightarrow H_{0}(\calT) \oplus H_{0}(T) \longrightarrow H_{0}(\calT') \longrightarrow 0 
    %     \end{CD}
    % \end{align*}
\end{remark}

@@

\section{Extension operator on finite element stars}\label{section:extension}

\section{Constructive estimate of Poincar\'e--Friedrichs constants}\label{section:poincarefriedrichs}
